### 2018-10-28 paper/50_cases.tex
intro

\section{Case Study} \label{cases_study}

\subsection{Abstract Compilation a la Staging}

\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation as a new
implementation technique for abstract interpretation based static analysis \cite{Boucher:1996:ACN:647473.727587}.
Their work is esentially a partial evaluated CE (\textit{Control-Environment}) abstract
interpreter written in a big-step semantics for programs written in continuation-passing style.
The analysis it can perform is a monovariant control flow analysis, sepcifically, Shiver's original
formulation of 0-CFA \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
After the partial evaluation on input program, the \textit{control} part of abstract interpreter
is disappeared, instead, the compiled analysis is a closure (higher-order function) and
only takes an environment as argument,
and the overhead of traversing the abstract syntax tree of source program also has been eliminated.
Since the target language is written a continuation-passing style, so there is
no explict evaluation context component when constructing the abstract interpreter
(rather than a CEK machine).

In this section, we firstly review their work of closure generation,
and reproduce their result by adding stage annotations to 0-CFA.
Show that \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation can be understood as an instance of staged 
abstract interpreter.

Advantages: (almost?) no need to modify the analysis code; staging the store data structure

(Fianally, we should briefly summarize the difference between them:
staging vs partial evaluation; code generation; engineering effort; performance)

(Specific -> General vs General vs Specific)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Control Flow Analysis}

\subsubsection{context/path/flow-sensitive}

\subsubsection{Store-widening}

\subsubsection{Mixed Sensitivity}

using different $k$ for $k$-CFA

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Numerical Analysis in Imperative Languages} \label{cases_imp}

Beyond control-flow analysis, staging data structures used in the abstract interpreter, such as abstract domain.
It has been shown that specializing abstract domains with respect to the structure of analyzed program significantly 
improves the performance: a recent example is online decomposition of polyhedra \cite{DBLP:conf/popl/SinghPV17}.
In this section, we present a similar idea for interval abstract domain using an \textit{offline} fashion by staging.

\subsubsection{Scaling to Imperative Languages}

Assignments, Loops, Exceptions

\subsubsection{Specializing Interval Domains}

Intervals, Numerical Analysis

Options: Data-flow analysis, taint analysis