### 2018-11-02 paper/abstract.tex
attempt at revising abstract

It is well known that a staged interpreter is a compiler: specializing the interpreter to a given program
produces an equivalent program that runs faster.
It is even more widely known that an abstract interpreter is a program analyzer: tweaking the 
interpreter to run on a domain of abstract values produces a sound static analysis.

What happens when we combine these two ideas, and apply staging to an \emph{abstract} interpreter? 
We obtain a sound static analysis, specialized for a given program, that runs faster.
%
More surprisingly, we show that by applying the staged abstract interpreter to \textit{open} programs and considering the free variables as dynamic inputs, 
we obtain a modular analysis that generates sound partial analysis results which can be composed 
and reused later without losing precision, even though the original abstract interpreter is a whole-program 
analysis algorithm.

Based on the idea of staged abstract interpreters, we show several case studies, including 
\citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation of 0-CFA, pushdown
control-flow analysis with context/path/flow-sensitivity and store-widening, and a numerical
analysis on an imperative language.

We empirically evaluate the performance improvements on control-flow analysis of benchmark programs.
The results show an average speedup of Nx with staging for a monovariant analysis, and Mx for 
a polyvariant analysis.

% It is well known that a staged interpreter is a compiler, which provides performance improvement
% by specializing the interpreter to a given program.

% In this paper, we study \textit{abstract} interpreters combined with multi-stage programming, i.e., the 
% staged abstract interpreters. 
% By staging the abstract interpreter with respect to a program, we obtain a specialized analysis that runs faster.
% By applying the staged abstract interpreter with \textit{open} programs and considering the free variables as dynamic inputs, 
% we obtain a modular analysis that generates sound partial analysis results which can be composed 
% and reused later without losing precision, though the original abstract interpreter is a whole-program 
% analysis algorithm.

% Using the idea of staged abstract interpreters, we show several case studies, including 
% \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation of  0-CFA, pushdown
% control-flow analysis with context/path/flow-sensitivity and store-widening, and a numerical
% analysis on an imperative language.

% We also empirically evaluate the improvement of performance on control-flow analysis of benchmark programs.
% The result shows an average speedup of Nx when staging to Scala for a monovariant analysis, and Mx for polyvariant analysis.