### 2018-11-10 paper/23_bg_stage.tex
add some stuff

\subsection{From Interpreter to Staged Interpreter} \label{stagedinterp}

\textbf{Multi-Stage Programming in LMS.}

\begin{lstlisting}
val power5 = new DslDriver[Int, Int] {
  def power(b: Rep[Int], x: Int): Rep[Int] =
    if (x == 0) 1 else b * power(b, x-1)
  def snippet(x: Rep[Int]): Rep[Int] = power(x, 5)
}
\end{lstlisting}

Lightweight modular staging (LMS) \todo{cite} is a library-based multi-stage programming framework 
that enables type-safe dynamic code generation. Different from the traditional approach of using
syntactic quasi-quotation, it solely relies on types to distinguish between binding times.

In the example shown above, \texttt{power} is a staged function that takes 2 arguments where 
\texttt{b} is declared as \texttt{Rep[Int]} meaning that \texttt{b} is a representation of type 
\texttt{Int} whose underlying value will be available at the next stage. Meanwhile the actual 
computation result of the \texttt{power} function  will be available at the next stage as well. 
Thus the return value of \texttt{power} is also of type \texttt{Rep[Int]}. The generated code for
\texttt{power5} consists of a function that takes only one argument that corresponds to its 
representation \texttt{b} from the previous stage.

\textbf{Staged Concrete Interpreter.}

We share the same concrete type instantiation, i.e. trait \texttt{Concrete}, with the unstaged version,
but reimplemented the staged version for the concrete operations.

\begin{lstlisting}
trait RepConcInterpOps extends Concrete with LMSOps {
  type R[+T] = Rep[T]
  implicit def valueTyp: Typ[Value]
  val $\rho$0: Rep[Env] = Map[Ident,Addr]()
  val $\sigma$0: Rep[Store] = Map[Addr,Value]()
  def get($\rho$: Rep[Env], x: Ident): Rep[Addr] = $\rho$(x)
  def put($\rho$: Rep[Env], x: Ident, a: Rep[Addr]): Rep[Env] = 
    $\rho$ + (unit(x) -> a)
  def get($\sigma$: Rep[Store], a: Rep[Addr]): Rep[Value] = $\sigma$(a)
  def put($\sigma$: Rep[Store], a: Rep[Addr], v: Rep[Value]): Rep[Store] = 
    $\sigma$ + (a -> v)
  def alloc($\sigma$: Rep[Store], x: Ident): Rep[Addr] = $\sigma$.size + 1
  def close(ev: EvalFun)($\lambda$: Lam, $\rho$: Rep[Env]): Rep[Value] = {
    val Lam(x, e) = $\lambda$
    val f: Rep[(Value,Store)]=>Rep[(Value,Store)] = 
      (as: Rep[(Value, Store)]) => {
        val arg = as._1; val $\sigma$ = as._2
        val $\alpha$ = alloc($\sigma$, x)
        ev(e, put($\rho$, x, $\alpha$), put($\sigma$, $\alpha$, arg))
      }
    unchecked("CompiledClo(", fun(f), ",", $\lambda$, ",", $\rho$, ")")
  }
  def num(i: Lit): Rep[Value] = {
    unchecked("NumV(", i.i, ")")
  }
  def branch0(cnd: Rep[Value], thn: => Ans, els: => Ans): Ans = {
    val i = unchecked[Int](cnd, ".asInstanceOf[NumV].i")
    (if (i == 0) thn else els).asInstanceOf[Rep[(Value,Store)]] //FIXME: Why?
  }
  def prim_eval(op: Symbol, v1: Rep[Value], v2: Rep[Value]): Rep[Value] = {
    val v1i = unchecked(v1, ".asInstanceOf[NumV].i")
    val v2i = unchecked(v2, ".asInstanceOf[NumV].i")
    unchecked("NumV(", v1i, op.toString.drop(1), v2i, ")")
  }
}
\end{lstlisting}

\textbf{Code Generation.}

code generation, runtime code