### 2018-11-11 paper/23_bg_stage.tex

\subsection{From Interpreter to Staged Interpreter} \label{stagedinterp}

\paragraph{Multi-Stage Programming in LMS.}
Lightweight modular staging (LMS) \cite{DBLP:conf/gpce/RompfO10} is a multi-stage programming framework implemented 
as a Scala library that enables dynamic code generation in a type-safe manner.
Different from the syntactic approach such as MetaOCaml \cite{DBLP:conf/flops/Kiselyov14} that uses quotations, 
LMS distinguishes binding-time solely based on types.
LMS provides a type constructor @Rep[T]@ where @T@ can be an arbitrary type, indicating an expression will be 
known at next stage. All operations act on a @Rep@ expression will be resulted in the generated code. However,
we need to provide intermediate representation and code generation support for type @T@ to LMS. 
Fortunately, the LMS framework already provides
such support for primitive types and commonly used data structure such as arrays and maps; implementing such support
for custom class is also straightforward. Let's go back and see how LMS can be used to specialize the power function
we mentioned in Section 1.

\begin{lstlisting}
val power5 = new DslDriver[Int, Int] {
  def power(b: Rep[Int], x: Int): Rep[Int] =
    if (x == 0) 1 else b * power(b, x-1)
  def snippet(x: Rep[Int]): Rep[Int] = power(x, 5)
}
\end{lstlisting}

In the code shown above, \texttt{power} takes 2 arguments where \texttt{b} is declared as \texttt{Rep[Int]} type 
meaning that \texttt{b} is a representation of \texttt{Int} but whose value will be available at the next stage. 
Meanwhile, the result of the @power@ function will be available at the next stage as will, thus the return value 
of \texttt{power} is also of type \texttt{Rep[Int]}. 
Then we specialize @power@ in function @snippet@ by providing @x@ to @5@.
The generated code for \texttt{power5} is a function that takes only one argument which corresponds to @b@,
and body of this function multiples @b@ five times.

\paragraph{Staged Concrete Interpreter.} In the staging part, we share the same concrete type instantiation, 
i.e. trait \texttt{Concrete}, but we reimplement the staged version for the concrete operations.
Meanwhile, @RepConcInterpOps@ also extends from @LMSOps@ which provides necessary supports for the 
staged version of primitive types and data structures.

\begin{lstlisting}
trait RepConcInterpOps extends Concrete with LMSOps {
  type R[+T] = Rep[T]
  val $\rho$0: Rep[Env] = Map[Ident,Addr]()
  val $\sigma$0: Rep[Store] = Map[Addr,Value]()
  def get($\rho$: Rep[Env], x: Ident): Rep[Addr] = $\rho$(x)
  def put($\rho$: Rep[Env], x: Ident, a: Rep[Addr]): Rep[Env] = 
    $\rho$ + (unit(x) -> a)
  def get($\sigma$: Rep[Store], a: Rep[Addr]): Rep[Value] = $\sigma$(a)
  def put($\sigma$: Rep[Store], a: Rep[Addr], v: Rep[Value]): Rep[Store] = 
    $\sigma$ + (a -> v)
  def alloc($\sigma$: Rep[Store], x: Ident): Rep[Addr] = $\sigma$.size + 1
  def num(i: Lit): Rep[Value] = NumV(i)
  def branch0(cnd: Rep[Value], thn: => Ans, els: => Ans): Ans = {
    //FIXME: the cast is ugly, patten match?
    val i = cnd.asInstanceOf[Rep[NumV]].i
    if (i == 0) thn else els
  }
  def prim_eval(op: Symbol, 
    v1: Rep[Value], v2: Rep[Value]): Rep[Value] = op match {
      case '+ => v1.asInstanceOf[Rep[NumV]].i + 
                 v2.asInstanceOf[Rep[NumV]].i
      ...
    }
  ...
}
\end{lstlisting}

The first notable change is the abstract type member @R@ is assigned to be @Rep@, and accordingly the affected 
types such as @Env@ and @Store@ become @Rep[Env]@ and @Rep[Store]@. \todo{NumV}
The interesting thing is how we handle closures and function applications when staging is involved.
In the unstaged version, function @close@ is used to lift a literal lambda term to a closure, 
however, what we desire is the a \textit{compiled} closure, i.e., the syntactic term of the body expression should be
eliminated and specialized.
The specialization of the interpreter with respect to the body of the lambda term proceed under the
assumptions that the argument and latest store will be provided later. \note{to denotational style?}
We will see this is an important observation that enables us to specialize an abstract interpreter modularly.
In the end of @close@, we create a @CompiledClo@ object that contains a staged function, which both will be
exhibited in the generated code.

\begin{lstlisting}
def close(ev: EvalFun)($\lambda$: Lam, $\rho$: Rep[Env]): Rep[Value] = {
  val Lam(x, e) = $\lambda$
  val f: Rep[(Value,Store)]=>Rep[(Value,Store)] = { 
    case (arg: Rep[Value],$\sigma$: Rep[Store]) =>
      val $\alpha$ = alloc($\sigma$, x)
      ev(e, put($\rho$, x, $\alpha$), put($\sigma$, $\alpha$, arg)) 
  }
  CompiledClo(fun(f))
}
\end{lstlisting}

Then, @apply_closure@ takes three arguments the function value, argument value and the latest store, 
which are both of @Rep@ type. It means basically these is nothing we can do at current stage --- 
so we create a new node @ApplyClosure@ that contains these arguments in the intermediate representation graph,
and in the code generation phase we can emit the code that does the application. \todo{reflectEffect}.

\begin{lstlisting}
case class ApplyClosure(f: Rep[Value], 
  arg: Rep[Value], $\sigma$: Rep[Store]) extends Def[(Value, Store)]
def apply_closure(ev: EvalFun)
  (f: Rep[Value], arg: Rep[Value], $\sigma$: Rep[Store]): Ans = {
    reflectEffect(ApplyClosure(f, arg, $\sigma$))
  }
\end{lstlisting}

\paragraph{Code Generation.}
When generating code for the next stage, values like @CompiledClo@ and @NumV@ are kept as what it is because
they are intended to be existed in the next stage; however, we need to treat @ApplyClosure@ a little bit 
differently.

\begin{lstlisting}
  case ApplyClosure(f, arg, $\sigma$) => 
    emitValDef(sym, quote(f) + 
                    ".asInstanceOf[CompiledClo].f(" + 
                    quote(arg) + "," + 
                    quote($\sigma$) + ")")
\end{lstlisting}

Concretely, as shown in the code above, when an @ApplyClosure@ is matched when emitting code, we just 
emit a function application where the function value is extracted from the @CompiledClo@
(we assume the object program is well-typed so that we can safely cast it to @CompiledClo@), 
and argument and store are the rest fields in @ApplyClosure@.