### 2018-11-12 paper/30_sai.tex

\section{Staging Abstract Interpreters} \label{sai}

In the previous section, we have seen an unstaged abstract interpreter and a staged concrete interpreter, 
now we begin describing the implementation of their confluence -- a staged abstract interpreter. 
Unsurprisingly, the staged abstract interpreter has the same abstract semantics with the unstaged version
we presented in Section~\ref{unstaged_abs}.

\subsection{From Abstract Interpreter to Staged Abstract Interpreter}

In this section, we present a principled approach to derive staged abstract interpreter
from its unstaged version. One guiding principle of our approach is the code of abstract semantics
and the code that optimizes should be separated. Thus it is also an advantage of doing staging 
for abstract interpreters: the designer of analysis has no need to rewrite the analysis, and 
the performance improvement comes almost for free, without any sacrifice of soundness or precision.

\paragraph{Staged Lattices.} The lattice definition we introduced in Section~\ref{unstaged_abs}
leaves space to be stage polymorphic, now we still use powerset as an example to present its 
staged version.

\begin{lstlisting}
trait RepLattice[A] extends Lattice[A, Rep]
implicit def RepSetLattice[T:Typ]: RepLattice[Set[T]] = 
  new RepLattice[Set[T]] {
    lazy val bot: Rep[Set[T]] = Set[T]()
    lazy val top: Rep[Set[T]] = throw new NotImplementedError()
    def $\sqsubseteq$(l1: Rep[Set[T]], l2: Rep[Set[T]]): 
      Rep[Boolean] = l1 subsetOf l2
    def $\sqcup$(l1: Rep[Set[T]], l2: Rep[Set[T]]): 
      Rep[Set[T]] = l1 union l2
    def $\sqcap$(l1: Rep[Set[T]], l2: Rep[Set[T]]): 
      Rep[Set[T]] = l1 intersect l2
  }
\end{lstlisting}

The type parameter @T:Typ@ of @RepLattice@ requires the element of sets
is also can be staged, otherwise without knowing how to stage the elements in the set
we can not stage the set as well.
The methods defined operate on type @Rep[Set[T]]@, thus the underlying implementation
such as @union@ and @intersect@ will be mapped to a node in IR graph and emitted in 
the generated code. Again, other structures such as maps and tuples are also implemented
in a similar way.

\paragraph{Staged Abstract Semantics.}

\begin{lstlisting}
trait RepAbsInterpOps extends Abstract with LMSOps {
  type R[+T] = Rep[T]
  val $\rho$0: Rep[Env] = Map[Ident, Addr]()
  val $\sigma$0: Rep[Store] = Map[Addr, Value]()
  def get($\rho$: Rep[Env], x: Ident): Rep[Addr] = $\rho$(x)
  def put($\rho$: Rep[Env], x: Ident, a: Rep[Addr]): 
    Rep[Env] = $\rho$ + (unit(x) -> a)
  def get($\sigma$: Rep[Store], a: Rep[Addr]): Rep[Value] = 
    $\sigma$.getOrElse(a, RepLattice[Value].bot)
  def put($\sigma$: Rep[Store], a: Rep[Addr], v: Rep[Value]): 
    Rep[Store] = $\sigma$ + (a -> RepLattice[Value].$\sqcup$(v, get($\sigma$, a)))
  def alloc($\sigma$: Rep[Store], x: Ident): Rep[Addr] = Addr(x)
  def num(i: Lit): Rep[Value] = Set(NumV())
  def close(ev: EvalFun)($\lambda$: Lam, $\rho$: Rep[Env]): Rep[Value] = {
    val Lam(x, e) = $\lambda$
    val f: Rep[(Value, Store)]=>Rep[(Value,Store)] = 
      (as: Rep[(Value,Store)]) => {
        val args = as._1; val $\sigma$ = as._2; val $\alpha$ = alloc($\sigma$, x)
        ev(e, put($\rho$, x, $\alpha$), put($\sigma$, $\alpha$, args))
      }
    Set[AbsValue](CompiledClo(fun(f)))
  }
  def branch0(cnd: Rep[Value], thn: => Ans, els: => Ans): Ans =
    thn $\sqcup$ els
  def prim_eval(op: Symbol, 
                v1: Rep[Value], v2: Rep[Value]): Rep[Value] = 
    Set(NumV())
  ...
}
\end{lstlisting}

\paragraph{Staged Fixpoint Iteration.} staged fixed point iteration with memoization

\paragraph{Specialized Data Structures.}

\subsection{Discussion}
what not to do when writing an abstract interpreter for staging?
what kind of abstract interpreter is friendly for staging/specialization?

Stage an abstract abstract machine yields another small-step abstract machine

Big-step vs small-step

direct-style vs CPS

Operational vs denotational

General insights for staging other static analysis algorithm.

\cite{10.1007/3-540-61580-6_11}

Jones-optimal specialization?
Futamura projection?

\subsection{Lattice-theoretic Interpretation}