### 2018-11-16 paper/70_related.tex

\section{Related Work}

\paragraph{Optimizing Static Analysis Through Specialization}
The idea in this paper is closely inspired by abstract compilation \cite{Boucher:1996:ACN:647473.727587}.
\citeauthor{Boucher:1996:ACN:647473.727587} presented abstract compilation
techniques as an efficient implementation of the monovariant flow analysis 
(\textit{0}-CFA) for programs written in continuation-passing style. 
The key idea is to remove the interpretation overhead on traversing the syntax 
tree by partial evaluation. Specifically, they proposed two similar kinds of abstract compilation techniques.
The first one is to generate specialized analysis as a textual program, which then
can be loaded and executed by \texttt{eval} or other similar mechanisms. 
The second one is to use closures, i.e., functions that remember their environments, 
as a representation of specialized analysis. As we show in the case study,
compiling the analysis generates higer-order functions on-the-fly with respect to 
the analyzed program, then the generated closure can be applied immediately in the 
higher-order host language.

\citeauthor{Johnson:2013:OAA:2500365.2500604} adapt the idea of closure generation 
to optimize small-step abstract interpreter in state-transition style \cite{Johnson:2013:OAA:2500365.2500604}.
The analyzed program is firstly compiled to a intermediate representation called 
"abstract bytecode", which are actually higher-order functions, and then be executed 
on a abstract abstract machine for that IR.

\citet{damian1999partial} provides a formal treatment to abstract compilation and Shiver's CFA, 
as well as proofs to establish correctness and certified specialized analyzers \cite{damian1999partial}.
\citeauthor{amtoft1999partial} applyies partial evaluation for constraint-based 
control flow analysis \cite{amtoft1999partial}.
Split the analysis to be multiple stages is also studied other than control-flow analysis,
though the formulation may very different.
For example, \citeauthor{DBLP:conf/cgo/HardekopfL11} apply staging to flow-sensitive pointer 
analysis \cite{DBLP:conf/cgo/HardekopfL11}. The first stage is to analyze the program code to 
obtain a sparse representation, and then the second stage conducts the flow-sensitive analysis.

\citeauthor{DBLP:conf/popl/SinghPV17} optimizes polyhedra abstract domain through online decomposition \cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143}.
The idea is to decompose a large polyhedra into several smaller one; one abstract transformer
usually only depends a limit number of variables, then we just need to compute the resulting polyhera
on those smaller polyhedra that contains these related variables.
The decomposition can be considered as a specialization of abstract domain with respect to the program
static structure.

\paragraph{Abstract Interpreters}
Abstract interpretation was proposed as a semantic-based approach to construct sound static analysis by approximation \cite{DBLP:conf/popl/CousotC77}.
As semantic artifacts, the Abstracting Abstract Machines (AAM) \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10}
approach shows that abstract interpreters can be derived systematically from concrete semantic artifacts.
The big-step abstract interpreters we presented in this paper are also inspired by the AAM framework,
which are later adopted for big-step abstract interpreters \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800}.
\citet{DBLP:conf/cc/CousotC02} also proposed an abstract interpretation framework for modular analysis.
\citet{DBLP:conf/popl/CalcagnoDOY09} developed compositional shape analysis by using bi-abduction.
The big-step abstract interpreter we presented in this paper is inspired by \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800}.

\textbf{Meta-Programming}.

Partial evaluation. 
Reflexive language \cite{DBLP:conf/gpce/Asai14}
Collapsing Towers of Interpreters \cite{Amin:2017:CTI:3177123.3158140}

The Lightweight Modular Staging framework in Scala
Other notable implementations of MSP exist in ML family, for examples, MetaML \cite{DBLP:conf/pepm/TahaS97} 
and MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14}.
Compared with the LMS approach in Scala, MetaML/MetaOCaml use term-level annotations such as brackets, 
escape, and run. Notwithstanding, we use LMS in this paper, the idea of staging an abstract interpreter
still applies with other MSP implementations.