### 2018-11-16 paper/60_eval.tex
Run evaluation

\section{Evaluation} \label{evaluation}

\subsection{Benchmarks}

We have designed, implemented and evaluated a prototype abstract interpreter for Scheme.
Our prototype includes both staged and unstaged versions of the abstract interpreter for comparison 
purposes.

We ran our prototype abstract interpreter on a suite of benchmark scheme programs that are also used 
in \cite{Johnson:2013:OAA:2500365.2500604, ashley:practical}.

\subsection{Performance}

\subsubsection{Baselines}
Small-step AAM

Big-step ADI

OAAM with AC \cite{Boucher:1996:ACN:647473.727587, Johnson:2013:OAA:2500365.2500604}

Benchmarks and other reference implementations: \url{https://github.com/ilyasergey/reachability}.

\subsection{Compositionality}