### 2018-11-17 paper/20_bg.tex
revise

\newcommand{\TLang}{$L_\lambda$}

\section{Preliminaries}

In this section, we first describe the abstract syntax of the language we will analyze, then present a
generic interface for interpreters, and instantiate an unstaged concrete interpreter.
Then, starting from this unstaged concrete interpreter, we proceed in two directions:
by abstracting it we obtain an unstaged abstract interpreter (Section~\ref{unstaged_abs}); 
by staging it we obtain a staged concrete interpreter (Section~\ref{stagedinterp}).

\input{21_bg_interp}
\input{23_bg_stage}
\input{22_bg_abs}