### 2018-11-17 paper/50_cases.tex

\section{Case Study} \label{cases_study}

In Section~\ref{sai}, we have shown that staging an abstract interpreter is feasible and can be 
systematic after correctly identifying the binding-times, despite the fact that the abstract 
interpreter is intended to be imprecise and easy to implement. 
In this section, we conduct several case studies to show that this approach is also useful and
widely applicable to different analyses.

\subsection{Abstract Compilation a la Staging} \label{cs_ac}

\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation (AC) as a new
implementation technique for abstract interpretation based static analysis \cite{Boucher:1996:ACN:647473.727587}.
The idea is inspired by partial evaluation similar to the present paper -- the program can be known 
statically therefore the overhead of interpretation can be eliminated. 
In AC, the compiled analysis can be represented by either text or closures (higher-order functions);
the closures can be executed immediately however the textual program need to be compiled and loaded first.

Specifically, \citeauthor{Boucher:1996:ACN:647473.727587} show how to compile a monovariant control-flow 
analysis \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007} 
for continuation-passing style (CPS) programs. Since the analyzed program is written in CPS, the analyzer 
is essentially a big-step control-environment abstract interpreter.
Closure generation compiles the analysis as a closure which takes an environment as argument.
The overhead of traversing the abstract syntax tree of input program also has been eliminated.

In this section, we show that \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation can be 
understood and implemented as an instance of staging abstract interpreters.
We first revisit the original implementation of abstract compilation of 0-CFA,
and then reproduce their result by simply adding stage annotations.
The generated program of our approach improves approximately the same extent of speed,
but without changing a single line of the analyzer program (with the use of LMS).
However, closure generation requires more engineering effort, specifically a whole-program 
conversion on the analyzer. Moreover, as shown in Section~\ref{staged_ds}, 
our approach is able to not only remove the interpretive overhead, but also specialize 
the data structures used in the analysis, for example, the environment that maps variables to sets of lambda.

\begin{figure*}
  \centering
  \begin{subfigure}[h]{0.49\textwidth}
    \centering
    \begin{lstlisting}
type CompAnalysis = Store => Store
def compProgram(prog: Expr): CompAnalysis = compCall(prog)
def compCall(call: Expr): CompAnalysis = call match {
  case Letrec(bds, body) =>
    val C1 = compCall(body); val C2 = compArgs(bds.map(_.value))
    ($\sigma$: Store) => C1(C2($\sigma$.update(bds.map(_.name), 
       bds.map(b => Set(b.value.asInstanceOf[Lam])))))
  case App(f, args) =>
    val C1 = compApp(f, args); val C2 = compArgs(args)
    ($\sigma$: Store) => C1(C2($\sigma$))
}
def compApp(f: Expr, args: List[Expr]): CompAnalysis = 
  f match {
    case Var(x) => ($\sigma$: Store) => 
      analysisAbsApp($\sigma$.lookup(x), args, $\sigma$)
    case Op(_) => compArgs(args)
    case Lam(vars, body) =>
      val C = compCall(body)
      ($\sigma$: Store) => 
        C($\sigma$.update(vars, args.map(primEval(_, $\sigma$))))
  }
def compArgs(args: List[Expr]): CompAnalysis = args match {
  case Nil => ($\sigma$: Store) => $\sigma$
  case (arg@Lam(vars, body))::rest =>
    val C1 = compCall(body); val C2 = compArgs(rest)
    ($\sigma$: Store) => C2(C1($\sigma$))
  case _::rest => compArgs(rest)
}
  \end{lstlisting}
  \end{subfigure}
\hfill
  \begin{subfigure}[h]{0.49\textwidth}
    \centering
    \begin{lstlisting}
def analyzeProgram(prog: Expr, $\sigma$: Rep[Store]): Rep[Store] = 
  analyzeCall(prog, $\sigma$)
def analyzeCall(call: Expr, $\sigma$: Rep[Store]): Rep[Store] = 
  call match {
    case App(f, args) => analyzeApp(f, args, analyzeArgs(args, $\sigma$))
    case Letrec(bds, body) =>
      val $\sigma$_* = $\sigma$.update(bds.map(_.name), 
        bds.map(b => Set(b.value.asInstanceOf[Lam])))
      val $\sigma$_** = analyzeArgs(bds.map(_.value), $\sigma$_*)
      analyzeCall(body, $\sigma$_**)
  }
def analyzeApp(f: Expr, args: List[Expr], $\sigma$: Rep[Store]): Rep[Store] = 
  f match {
    case Var(x) => analyzeAbsApp(args, $\sigma$(x), $\sigma$)
    case Op(_) => analyzeArgs(args, $\sigma$)
    case Lam(vars, body) =>
      val $\sigma$_* = $\sigma$.update(vars, args.map(primEval(_, $\sigma$)))
      analyzeCall(body, $\sigma$_*)
  }
def analyzeArgs(args: List[Expr], $\sigma$: Rep[Store]): Rep[Store] = 
  args match {
    case Nil => $\sigma$
    case Lam(vars, body)::rest => analyzeArgs(rest, analyzeCall(body, $\sigma$))
    case _::rest => analyzeArgs(rest, $\sigma$)
  }
  \end{lstlisting}

  \end{subfigure}
  \caption{Comparison of AC (left) and SAI (right). Only core code are shown.}
  \label{compare_ac_sai}
\end{figure*}

\subsubsection{Closure Generation}

The analysis presented by \citeauthor{Boucher:1996:ACN:647473.727587} is 0-CFA for a CPS language consisting
of lambda terms, applications, @letrec@ and primtive operators. The analyses for different syntactic 
constructs are decomposed into different functions, such as @analyzeCall@ and @analyzeApp@.
The idea of closure generation is to rewrite these functions. Where previously they may take both static
arguments and dynamic arguments, after the rewrite only the static arguments is taken.
In this case, the static arguments are syntactic terms;
the dynamic arguments are stores.
After written in AC style, functions like @compCall@ (compiled version of @analyzeCall@) return a value of type
@CompAnalysis@, i.e., a closure that takes a store and returns a store.
The result of multiple calls on such functions, for example,
@compCall@ and @compArgs@, can be composed.
The generated closure only takes stores, because the input program
is specialized into the closure. The code is shown in Figure~\ref{compare_ac_sai} (left).

\subsubsection{Staged 0-CFA}

On the other side, our approach does exactly the same thing through staging: the syntactic terms are static, and 
stores are dynamic, therefore the generated code just looks-up and updates the store.
Figure~\ref{compare_ac_sai} (right) shows the code written with LMS. In fact, the only changes are
the type of @Store@ is replaced with @Rep[Store]@ indicating that the values of type @Store@  will be known
at the next stage. Indeed, additional engineering efforts are required to make this happen, 
including: the staged version @Map@ which is already included in LMS; implicit @lift@ function that transform a
current-stage constant value to next stage; next-stage representation of the syntactic terms, i.e., proper @toString@
functions of AST structs. We consider these efforts relatively small, and they do not interfere the actual analysis
we desire.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Control-flow Analysis} \label{cfa}

The target language we presented in Section~\ref{bg_lang} is essentially a higher-order functional language.
One fundamental analysis task for functional programs is control-flow analysis, i,e., determining
which functions will possibly be applied at each call-site. 
The abstract interpreter we used in Section~\ref{unstaged_abs} and Section~\ref{sai} is a store-widened 
0-CFA-like abstraction, moreover it is also a pushdown control-flow analysis; 
in the last section, we also reviewed AC with finite-state 0-CFA.
In this section, based on the existing staged abstract interpreter, we further
develop the staging techniques with control-flow analyses, including recoverying a more precise store model and
adding context-sensitivity to the analysis.
%, and using staging as an implementation stategy for mixed sensitivity.

\subsubsection{A More Precise Store Model}

The store-widened analysis we implemented in Section~\ref{sai} uses a single store to approximate
the runtime store. It can be efficently computed in polynomial time together with 0-CFA-like abstraction, 
but sometimes we may desire a 
more precise result that distinguishes the final values and stores for different closure targets. 
To achieve this goal,
we need to tweak our abstract interpreter and type instantiation. The answer type @Ans@ is changed to
a set of @VS@s where a @VS@ is a pair of sets of abstract values (such as closures or abstract numbers)
and a store.

\begin{lstlisting}
type VS = (Set[AbsValue], Store)
type Ans = R[Set[VS]]
\end{lstlisting}

Note that the type @Ans@ uses our stage polymorphic type @R@, meaning that under staging the type @Ans@
represents a next stage value. Then, our generic interpreter is also changed when handling function 
applications.

\begin{lstlisting}
case App(e1, e2) =>
  val e1ans = ev(e1, $\rho$, $\sigma$)
  val e1vs  = choices(e1ans)
  val e2ans = ev(e2, $\rho$, e1vs.$\sigma$)
  val e2vs  = choices(e2ans)
  apply_closure(ev)(e1vs.v, e2vs.v, e2vs.$\sigma$)
\end{lstlisting}

The idea to handle the application is to explore all possible closures from @e1@ and meanwhile
use the latest store.
What @choices@ does is similar to McCarthy's @amb@ operator ~\cite{MCCARTHY196333}: it non-deterministically returns an
element of type @VS@ from its argument, e.g., @e1ans@, to its right-hand side receiver, i.e., @e1vs@.
The function @choices@ internally uses the delimited control operator @shift@ to capture the continuation,
which is the code block after its call site. 
This allows us to perform nesting depth-first evaluation for function application while still writing 
the program in direct-style \cite{Wei:2018:RAA:3243631.3236800}. 
Again, we can stage this part as we did for the naive abstract interpreter.

\subsubsection{Context-Sensitivity}

We add k-CFA-like context-sensitivity to the analysis by introducing an abstract timestamp, 
whose concrete instantiation is a finite list of expressions that track $k$ recent calling contexts.
The definition of abstract addresses is changed to a tuple of identifiers and the time it get allocated,
meaning that this address points to some values under such calling context.
If $k$ is 0, we obtain a monovariant analysis as demonstrated before; if $k > 0$, we obtain
a family of analysis with increasing precision.

\begin{lstlisting}
type Time = List[Expr]
\end{lstlisting}

Every time when we call the @eval@ function, we refresh the timestamp by calling function @tick@,
which returns a new timestamp. Here we adopt a $k$-CFA-like allocation strategy, therefore
the @tick@ function can be implemented as appending the current expression being evaluated to the existing
calling context, and then taking the first $k$ elements from the list.

\begin{lstlisting}
def tick(e: Expr, $\tau$: R[Time]): R[Time] = (e :: $\tau$).takewhile
def eval(ev: EvalFun)
  (e: Expr, $\rho$: R[Env], $\sigma$: R[Store], $\tau$: R[Time]): Ans = {
    val $\tau$_* = tick(e, $\tau$)
    e match {
      case Lit(i) => ...
      ...
    }
  }
\end{lstlisting}

Accordingly, the type of return value is accompanied by the timestamp. For a recursive call of @ev@,
which would also return the latest timestamp, and that timestamp will be used for the evaluation afterward.

\begin{lstlisting}
type VST = (Value, Store, Time)
type Ans = R[Set[VST]]
\end{lstlisting}

Using other allocation strategies to achieve different sensitivities is also possible \cite{DBLP:conf/icfp/Gilray0M16} 
and can be staged under our framework.

%\subsubsection{Mixed Sensitivity}

%using different $k$ for $k$-CFA

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Numerical Analysis in Imperative Languages} \label{cases_imp}

Now we consider in a first-order imperative language, we may care more about the data-flow because the control-flow
is relatively easy to obtain. In this section, we show the staging of other abstract domains, particularly an 
interval domain for numbers.
It has been shown that specializing abstract domains with respect to the structure of analyzed program significantly 
improves the performance: a recent example is online decomposition of polyhedra \cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143}.
In this section, we first show how to support imperative language features in the generic abstract interpreter.
Then we present a similar idea for the interval domain and show that the specialization is feasible by staging systematically.

\subsubsection{Scaling to Imperative Languages}

To evaluates an assignment, we first evaluate its right-hand side, and then put the value
into the slot where the address of @v@ points to.
For simplicity, we elect to make the value of the assignment be an @void()@ value.

\begin{lstlisting}
case Assign(x, e) =>
  val (v, $\sigma$_*) = ev(e, $\rho$, $\sigma$)
  (void(), put($\sigma$_*, get($\rho$, x), v))
\end{lstlisting}

To evaluates a @while@ loop statement, we evaluate the condition first.
Then similar to how we treat @branch0@, we have a generic @branch@ function but works on boolean
values. For the @true@ branch, we recursively call @ev@ on a newly constructed expression
@Seq(e, While(t, e))@ meaning that first evaluates @e@ and the repeat the loop.
Otherwise, for the other branch, we simply return a void value and current store.

\begin{lstlisting}
case While(t, e) =>
  val (tv, t$\sigma$) = ev(t, $\rho$, $\sigma$)
  branch(tv, ev(Seq(e, While(t, e)), $\rho$, t$\sigma$), (void(), t$\sigma$))
\end{lstlisting}

\subsubsection{Staged Interval}

\begin{lstlisting}
case class Interval(lb: Rep[Double], ub: Rep[Double]) {
  def +(that: Interval): Interval = that match {
    case Interval(lb_, ub_) => Interval(lb + lb_, ub + ub_)
  }
  def -(that: Interval): Interval = that match {
    case Interval(lb_, ub_) => Interval(lb - lb_, ub_ - ub) 
  }
  ...
}
\end{lstlisting}