### 2018-11-17 paper/80_concl.tex

\section{Conclusion}

In this paper, we show the feasibility of using multi-staging programming as a systematic approach 
to optimize abstract interpreters, while minimally modifying the abstract interpreter and keeping
the analysis sound.
We presented an end-to-end design and implementation
of an staged abstract interpreter for a small higher-order stateful language, which 
models the core part of modern programming languages.
The design that uses abstracting definitional interpreters captures the commonality between
abstract interpreters and concrete interpreters; the implementation that uses staging polymorphism
captures the commonality between unstage programs and staged programs.