### 2018-11-17 paper/60_eval.tex
revise

\section{Evaluation} \label{evaluation}

To evaluate the performance can be improved through staging, we use the abstract
interpreter demonstrated in the paper to analyze some Scheme programs. We implement
a desugaring transformation for a large subset of scheme that transforms into the
small language we used in the paper.
We use the same generic interface with a 0-CFA-like abstraction, and the unstage abstract
interpreter forms the baseline.

All of our evaluation benchmarks were performed on an Ubuntu 16.04 LTS (Linux kernel 4.4.0)
machine with 4 Intel Xeon Platinum 8168 CPU at 2.7GHz and 3 TiB of RAM.
Although the machine has 96 cores and 192 threads in total,
the abstract interpreters only use one thread to run all the benchmark programs.

\subsection{Benchmarks}

Our evaluation is based on a suite of benchmark tests. Considering that our abstract 
interpreters are implemented in Scala which will be affected by JVM warm up times, programs
tend to run faster after a number of runs. In order to minimize the influence of JVM, 
we ran all experiments 10 times and take the statistical mean and median values of 
the running times.

%on a suite of benchmark scheme programs that are also used in \cite{Johnson:2013:OAA:2500365.2500604, ashley:practical}. 
We used the following benchmark programs:
\begin{itemize}
    \item \textbf{fib:} a program that recursively calculates the $n$-th fibonacci number in
        exponential time.
    \item \textbf{church:} a program introduced by \todo{cite Dimitrios Vardoulakis and Olin Shivers. CFA2: a Context-Free approach
        to Control-Flow analysis. Logical Methods in Computer Science,
        7(2), 2011.} to test distributivity of multiplication for numbers in Church Encoding \todo{ cite }
\end{itemize}

\subsection{Performance}

For each benchmark program, we ran our prototype abstract interpreters, both unstaged and staged.
\todo{Figure show data.}

Our evaluations have shown that the staged abstract interpreter performs well in general.
Particularly in some interesting cases that are considered to be the worst case scenario
in the control-flow analysis, the staged version outperforms the unstaged abstract 
interpreter by a wide margin.

However, there also exists some programs where the unstaged abstract interpreter yields 
results faster than the unstaged interpreter.