### 2019-03-23 paper/20_bg.tex
revise

\newcommand{\TLang}{$L_\lambda$}

\section{Preliminaries}

In this section, we first describe the abstract syntax of the language we will
analyze, then present the generic interpreter shared among the four different
semantics, after which, we instantiate the interpreter to the concrete one. Then we
will proceed in two directions: by staging it we obtain a staged concrete interpreter
(Section~\ref{stagedinterp}); by abstracting it we obtain an unstaged abstract
interpreter (Section~\ref{unstaged_abs}).

We choose to use Scala and monad to demonstrate the idea, but the approach
is not restricted to our choice. One can use imperative style and other MSP languages 
(such as MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14} 
or Template Haskell \cite{Sheard:2002:TMH:636517.636528} \todo{cite the recent one})
to construct such staged abstract interpreters.

\input{21_bg_interp}
\input{22_bg_stage}
\input{23_bg_abs}