### 2019-08-31 paper/90_conc.tex

\section{Conclusion}

In this paper, we present a simple and elegant framework that extends the first
Futamura projection to abstract interpreters.
To realize it, we design a generic interpreter that abstracts over both
binding-time and value domains; by instantiating it with staged monads for
abstract interpretation, the staged abstract interpreter can be readily derived.
The meta-programming approach brings new insight and implementation techniques
to optimize static analyzers. The optimization can be realized in a completely
orthogonal way to the interpreter and abstract semantics.
In our empirical evaluation, staging improves performance by an order of magnitude
in terms of running time.