### 2019-09-03 paper/90_conc.tex

\section{Conclusion}

We present a simple and elegant framework that extends the first Futamura
projection to abstract interpreters.
To realize it, we design a generic interpreter that abstracts over both
binding-times and value domains; by instantiating it with staged monads for
abstract interpretation, the staged abstract interpreter can be readily derived.
The meta-programming approach brings new insights and implementation techniques
to optimize static analyzers. Optimizations can be achieved orthogonally to 
the interpreter and abstract semantics.  In our empirical evaluation, staging
improves performance by an order of magnitude in terms of running time.