### 2019-09-06 paper/30_stageconc.tex

\section{From Interpreters to Staged Interpreters} \label{stagedinterp}

In this section, based on the generic interpreter and concrete components we
presented in Section \ref{prelim}, we show how to stage the concrete
interpreter by changing the monad type and refactoring primitive
operations. We begin by briefly introducing the Lightweight Modular Staging
framework in Scala, and then replay the same steps from the unstaged
counterpart. Finally, we briefly describe the code generation.

\subsection{Multi-Stage Programming with LMS}

Lightweight modular staging (LMS) \cite{DBLP:conf/gpce/RompfO10} is a
multi-stage programming framework implemented as a Scala library. LMS enables
runtime code generation in a type-safe manner. Different from the approach of
MetaML/MetaOCaml \cite{DBLP:conf/flops/Kiselyov14, DBLP:conf/gpce/CalcagnoTHL03} which
use syntactic quotations, LMS distinguishes binding-time based on types. 
LMS provides a type constructor @Rep[T]@ where @T@ can be
an arbitrary type, indicating a value of type @T@ will be known at the future stage.
Operations acting on @Rep[T]@ expressions will be residualized as generated
code.

A classic example for introducing multi-stage programming is the power function
that computes $b^x$. Usually, it is implemented as a recursive function:
\begin{lstlisting}
  def power(b: Int, x: Int): Int = if (x == 0) 1 else b * power(b, x - 1)
\end{lstlisting}

If @x@ is a value known at the current stage, we may specialize the power function to
the value @x@ -- by unrolling the recursive calls. In LMS, this is fulfilled by
first adding the @Rep@ type annotation to variables known at the next stage. In this case,
@b : Rep[Int]@ is known later, and @x : Int@ is known currently, as shown in the
below code recipe.
The way we use this staged @power@ function is to create a @DslDriver@ and override the
@snippet@ method that gives the currently know value @x@ (5 in this example).
\begin{lstlisting}
  new DslDriver[Int, Int] {
    def power(b: Rep[Int], x: Int): Rep[Int] = if (x == 0) 1 else b * power(b, x-1)
    def snippet(b: Rep[Int]): Rep[Int] = power(b, 5) // specialize the power to b^5
  }
\end{lstlisting}

The LMS framework provides staging support for primitive data types such as
@Int@ and @Double@, and commonly-used data structures such as lists and maps.
The idea behind the framework is to construct a sea-of-node intermediate
representation (IR) for the next-stage program at the current stage
\cite{DBLP:conf/birthday/Rompf16}. For convenience, the conversion from
expressions of type @Rep[_]@ to their IR is done by using the implicit design
pattern. As we will see later, implementing staging and code generation
support for user-defined classes is also straightforward.

\subsection{Staged Concrete Semantics}

To implement the staged concrete interpreter, we replay the steps from the
instantiation of the unstaged concrete interpreter in Section \ref{unstaged_conc}.
However, now we use the @Rep@ type to annotate the value domains, environments and
stores, and redefine the staged version of monads and primitive operations.

\paragraph{Staged Monads}
We use the same structure of monad stack in the unstaged interpreter: a reader
monad with a state monad. The difference here is, now the monads operate on
staged values.  For brevity, we call them \textit{staged monads}. In the code,
we also use a @Rep@ prefix on the constructors and types to differentiate them.
But it is important to note that instances of the case class @ReaderT@/@StateT@
are not staged; also the monadic computation, such as functions of @R => M[A]@,
are not staged. Instead, the internal data that these monads operate on are
staged, i.e., @R@ and @A@ in the reader monad, @S@ and @A@ in the state monad.
The following code snippet shows the idea. We use \hl{light gray} to highlight
where @Rep@ type is added:
\begin{lstlisting}[escapechar=!]
  case class RepReaderT[M[_]: RepMonad, R, A](run: !\hl{Rep[R]}! => M[A]) {
    def flatMap[B](f: !\hl{Rep[A]}! => ReaderT[M, R, B]): RepReaderT[M, R, B] =
      RepReaderT(r => RepMonad[M].flatMap(run(r))(a => f(a).run(r))); ... }
  case class RepStateT[M[_]: RepMonad, S, A](run: !\hl{Rep[S]}! => M[(A, S)]) {
    def flatMap[B](f: !\hl{Rep[A]}! => StateT[M, S, B]): RepStateT[M, S, B] =
      RepStateT(s => RepMonad[M].flatMap(run(s))(as => f(as._1).run(as._2))); ... }
\end{lstlisting}

The function @f@ passed to @flatMap@ is also a function value known at the current stage.
Therefore all invocations of @flatMap@s can be reduced before generating code.
The fact that we only stage data but not monadic computation or monadic
values is the reason that we can peel of the monad stack in the generated code.
Now the type @AnsM[_]@ is instantiated as the same structure as before, but using the
@Rep@ versions:
\begin{lstlisting}
  type R[T] = Rep[T]
  type AnsM[T] = RepReaderT[RepStateT[RepIdM, Store, ?], Env, T]
\end{lstlisting}

Readers may notice that the conversion between unstaged monads and staged monads
is merely changing the type of unstaged data to staged data, which in fact can
be achieved without modifying the implementation of monads. This is true so far,
but as we will see, it is not straightforward for nondeterminism monad
(@SetT@) in the abstract interpreter. Because not only the elements in the set are
staged, but the whole set is also staged, and we have no knowledge about how
many elements in the set. In this section, we explicitly distinguish the two
versions of monadic interfaces; later we will instantiate the staged set monad
by manually fusing a fragment of the monad stack.

\paragraph{Primitive Operations} Most of the primitive operations can be easily
translated to their staged versions -- we just need to change the types.
As we mentioned before, we will take more care on how functions and
applications are handled.  One problem here is what should we do for
$\lambda$ terms when staging? We cannot directly create a next-stage
defunctionalized closure for it, because that means we still need to
interpret over these closure at the next stage.  The desired goal is to
eliminate the interpretation overhead by staging, so the right thing is to
compile the lambda term with its environment to a next-stage Scala function. To
do this, we need to recursively call the interpreter function at current stage,
i.e., the @ev@ argument passed to @close@.  The following code implements the
idea:
\begin{lstlisting}
  type ValSt = (Value, Store)
  def emit_compiled_clo(f: (Rep[Value], Rep[Store]) => Rep[ValSt], λ: Lam, ρ: Exp[Env]): Rep[Value]
  def close(ev: Expr => Ans)(λ: Lam, ρ: Rep[Env]): Rep[Value] = {
    val Lam(x, e) = λ
    val f: (Rep[Value], Rep[Store]) => Rep[ValSt] = { case (v: Rep[Value], σ: Rep[Store]) =>
        val α = alloc(σ, x)
        ev(e)(ρ + (unit(x) → α))(σ + (α → v))
    }; emit_compiled_clo(f, λ, ρ)
  }
\end{lstlisting}

In @close@, we first create a current-stage function @f@ that takes a
next-stage value and store, and returns staged values. Then we use
@emit_complied_clo@ to denote @f@ to a next-stage Scala function, represented
by the type @Rep[Value]@. Inside of @f@, we can access the evaluator via @ev@.
However, @ev@ produces a monadic value of type @AnsM@, which can only exist at
the current stage. To connect the current-stage monadic values and future-stage
grounded values, the evaluator @ev@ is invoked on the body expression @e@ to be
specialized w.r.t. @e@, and then we \textit{collapse} the monadic value of
@AnsM@ to grounded values by providing the prepared environment and store.

For function applications @ap_clo@, we have two \textit{next-stage} values @fun@
and @arg@. But what we can do on the values of @Rep@ type is limited. In fact, @fun@ does
not have an intensional application operation we can use directly. We only know
that it is a staged value produced by @close@, and represents a
next-stage Scala function. Therefore we can just generate a next-stage function
application, also in Scala. Since @ap_clo@ produces a monadic value of type
@Ans@, we can still use the monadic operation @get_store@ to obtain the latest
store object. Then @emit_ap_clo@ is used to generate a current-stage
representation of the future-stage application result, which is a pair of values
and stores @ValSt@. After getting that result @vs@, we use @put_store@ to
update the store and reify the result into the current-stage monadic value.
\begin{lstlisting}
  def emit_ap_clo(fun: Rep[Value], arg: Rep[Value], σ: Rep[Store]): Rep[ValSt]
  def ap_clo(ev: EvalFun)(fun: Rep[Value], arg: Rep[Value]): Ans = for {
    σ  <- get_store
    vs <- lift[ValSt](emit_ap_clo(fun, arg, σ))
    _  <- put_store(vs._2)
  } yield vs._1
\end{lstlisting}

Compared with the unstaged version, we can observe a shift of using the
evaluation function @ev@ from application-time to value-representation-time: in
the unstaged interpreter, invocations of @ev@ happens in @ap_clo@; while in the
staged interpreter, we eagerly apply @ev@ when denoting lambda terms to our
next-stage value domain.

\subsection{A Little Bit of Code Generation}

In the previous section, we merely show the types of @emit_compiled_clo@ and
@emit_ap_clo@ without concrete implementations. In this section, we briefly
discuss the IR node generated by them and sketch code generation.
We define the IR nodes @IRCompiledClo@ and @IRApClo@ using @case class@
extending from @Def[T]@. @Def[T]@ is a built-in type from the LMS framework,
representing next-stage value definitions of type @T@.
\begin{lstlisting}
  case class IRCompiledClo(f: Rep[(ValSt) => ValSt], λ: Lam, ρ: Rep[Env]) extends Def[Value]
  case class IRApClo(rator: Rep[Value], rand: Rep[Value], σ: Rep[Store]) extends Def[ValSt]
\end{lstlisting}

The IR nodes are manipulated by the LMS passes and finally generated to
next-stage Scala code. To generate code, LMS provides an @emitNode@ method for
programmers to control what is generated for each kind of nodes. Here we match
@IRCompiledClo@ and @IRApClo@ and generate definitions via @emitValDef@ for
them.  For compiled closures, we put the function @f@ and its accompanying
syntactic term and environment to its next-stage value representation
@CompiledClo@. For applications, we have the knowledge indicating that @fun@ is
a @CompiledClo@ at the stage, and it has a callable field @f@. Hence, we
generate a next-stage function application @f(arg, σ)@. The invocations of
@quote@s are used to generate code for the other next-stage values.
\begin{lstlisting}
override def emitNode(sym: Sym[Any], rhs: Def[Any]) = rhs match {
  case IRCompiledClo(f, λ, ρ) => emitValDef(sym, s"CompiledClo($\textdollar${quote(f)}, $\textdollar${quote(λ)}, $\textdollar${quote(ρ)})")
  case IRApClo(rator, rand, σ) => emitValDef(sym, s"$\textdollar$rator.f($\textdollar${quote(rand)}, $\textdollar${quote(σ)})")
  ...
}
\end{lstlisting}