### 2019-09-06 paper/30_stageconc.tex
revision

\section{From Interpreters to Staged Interpreters} \label{stagedinterp}

In this section, based on the generic interpreter and concrete components we
presented in Section \ref{prelim}, we show how to stage the concrete
interpreter by changing the monad type and refactoring primitive
operations. We begin by briefly introducing the Lightweight Modular Staging
framework in Scala, and then replay the same steps from the unstaged
counterpart. Finally, we briefly describe the code generation.

\subsection{Multi-Stage Programming with LMS}

Lightweight modular staging (LMS) \cite{DBLP:conf/gpce/RompfO10} is a
multi-stage programming framework implemented as a Scala library. LMS enables
runtime code generation in a type-safe manner. Different from the approach of
MetaML/MetaOCaml \cite{DBLP:conf/flops/Kiselyov14, DBLP:conf/gpce/CalcagnoTHL03} which
use syntactic quotations, LMS distinguishes binding-time based on types. 
LMS provides a type constructor @Rep[T]@ where @T@ can be
an arbitrary type, indicating a value of type @T@ will be known at the future stage.
Operations acting on @Rep[T]@ expressions will be residualized as generated
code.

A classic example for introducing multi-stage programming is the power function
that computes $b^x$. Usually, it is implemented as a recursive function:
\begin{lstlisting}
  def power(b: Int, x: Int): Int = if (x == 0) 1 else b * power(b, x - 1)
\end{lstlisting}

If @x@ is a value known at the current stage, we may specialize the power function to
the value @x@ -- by unrolling the recursive calls. In LMS, this is fulfilled by
first adding the @Rep@ type annotation to variables known at the next stage. In this case,
@b : Rep[Int]@ is known later, and @x : Int@ is known currently, as shown in the
below code snippet.
The way we use this staged @power@ function is to create a @DslDriver@ and override the
@snippet@ method that supplies 5 as the currently know value for @x@.
\begin{lstlisting}
  new DslDriver[Int, Int] {
    def power(b: Rep[Int], x: Int): Rep[Int] = if (x == 0) 1 else b * power(b, x-1)
    def snippet(b: Rep[Int]): Rep[Int] = power(b, 5) // specialize the power to b^5
  }
\end{lstlisting}

The LMS framework provides staging support for primitive data types such as
@Int@ and @Double@, and commonly-used data structures such as lists and maps.
The framework internally constructs a sea-of-nodes intermediate representation (IR)
for the next-stage program when executing the current stage \cite{DBLP:conf/birthday/Rompf16}.
For convenience, the conversion from expressions of type @Rep[_]@ to their IR
is done by using the implicit design pattern. As we will see later,
implementing staging and code generation support for user-defined classes is
also straightforward.

\subsection{Staged Concrete Semantics}

To implement the staged concrete interpreter, we replay the steps from the
instantiation of the unstaged concrete interpreter in Section \ref{unstaged_conc}.
However, now we use the @Rep@ type to annotate the value domains, environments and
stores, and redefine the staged version of monads and primitive operations.

\paragraph{Staged Monads}
We use the same monad stack structure as in the unstaged interpreter: a reader
monad with a state monad. The difference here is that, now the monads operate on
staged values.  For brevity, we call them \textit{staged monads}. In the code,
we also use a @Rep@ prefix on the constructors and types to differentiate them.
But it is important to note that instances of the case class @ReaderT@/@StateT@
are not staged, and that the monadic computation, such as functions of @R => M[A]@,
are also not staged. Instead, the internal data that these monads operate on are
staged, i.e., @R@ and @A@ in the reader monad, @S@ and @A@ in the state monad.
The following code snippet shows the idea. We use \hl{light gray} to highlight
where @Rep@ types are added:
\begin{lstlisting}[escapechar=!]
  case class RepReaderT[M[_]: RepMonad, R, A](run: !\hl{Rep[R]}! => M[A]) {
    def flatMap[B](f: !\hl{Rep[A]}! => ReaderT[M, R, B]): RepReaderT[M, R, B] =
      RepReaderT(r => RepMonad[M].flatMap(run(r))(a => f(a).run(r))); ... }
  case class RepStateT[M[_]: RepMonad, S, A](run: !\hl{Rep[S]}! => M[(A, S)]) {
    def flatMap[B](f: !\hl{Rep[A]}! => StateT[M, S, B]): RepStateT[M, S, B] =
      RepStateT(s => RepMonad[M].flatMap(run(s))(as => f(as._1).run(as._2))); ... }
\end{lstlisting}

The function @f@ passed to @flatMap@ is also a function value known at the current stage.
Therefore all invocations of @flatMap@ can be reduced before generating code.
The fact that we only stage data but not monadic computation or monadic
values is the reason that we can peel of the monad stack in the generated code.
Now the type @AnsM[_]@ is instantiated as the same structure as before, but using the
@Rep@ versions:
\begin{lstlisting}
  type R[T] = Rep[T]
  type AnsM[T] = RepReaderT[RepStateT[RepIdM, Store, ?], Env, T]
\end{lstlisting}

Readers may notice that the conversion between unstaged monads and staged monads
is merely changing the type of unstaged data to staged data, which in fact can
be achieved without modifying the implementation of monads. This is true so far,
but as we will see, it is not as straightforward for the nondeterminism monad
(@SetT@) in the abstract interpreter, because not only the elements in the set are
staged, but the whole set is also staged, and we have no knowledge about how
many elements in the set. In this section, we explicitly distinguish the two
versions of monadic interfaces. Later, we will instantiate the staged set monad
by manually fusing a fragment of the monad stack.

\paragraph{Primitive Operations} Most of the primitive operations can be easily
translated to their staged versions -- we just need to change the types.
As we mentioned before, we will take more care on how functions and
applications are handled.  One problem here is what should we do for
$\lambda$ terms when staging? We cannot directly create a next-stage
defunctionalized closure for it, because that means we still need to
interpret over these closure at the next stage.  The desired goal is to
eliminate the interpretation overhead by staging, so the right thing is to
compile the lambda term with its environment to a next-stage Scala function. To
do this, we need to recursively call the interpreter function at the current stage,
i.e., the @ev@ argument passed to @close@.  The following code implements the
idea:
\begin{lstlisting}
  type ValSt = (Value, Store)
  def emit_compiled_clo(f: (Rep[Value], Rep[Store]) => Rep[ValSt], λ: Lam, ρ: Exp[Env]): Rep[Value]
  def close(ev: Expr => Ans)(λ: Lam, ρ: Rep[Env]): Rep[Value] = {
    val Lam(x, e) = λ
    val f: (Rep[Value], Rep[Store]) => Rep[ValSt] = { case (v: Rep[Value], σ: Rep[Store]) =>
        val α = alloc(σ, x)
        ev(e)(ρ + (unit(x) → α))(σ + (α → v))
    }; emit_compiled_clo(f, λ, ρ)
  }
\end{lstlisting}

In @close@, we first create a current-stage function @f@ that takes a
next-stage value and store, and returns staged values. Then we use
@emit_complied_clo@ to denote @f@ to a next-stage Scala function, represented
by the type @Rep[Value]@. Inside of function @f@, we can access the evaluator via @ev@.
However, @ev@ produces a monadic value of type @AnsM@, which can only exist at
the current stage. To connect current-stage monadic values and future-stage
grounded values, the evaluator @ev@ is supplied with not only the body
expression @e@ to be specialized, but also the prepared environment and store.
By doing so, we \textit{collapse} the monadic value of @AnsM@ to grounded
values within staging; therefore in the future stage, there is no monads.

For function applications @ap_clo@, we have two \textit{next-stage} values @fun@
and @arg@. But what we can do with these values of type @Rep@ is limited. 
In fact, @fun@ does not have an intensional application operation we can use
directly. We only know that it is a staged value produced by @close@, and
it represents a next-stage Scala function. Therefore, we can only generate a
next-stage function application, also in Scala. @ap_clo@ produces a monadic
value of type @Ans@, so that we can still use the monadic operation @get_store@
to obtain the latest store. Then @emit_ap_clo@ takes the function and argument
values, and the latest store to generate a current-stage representation of the
future-stage application result, which is of type @Rep[ValSt]@, a pair of
values and stores.  After obtaining the result @vs@, we use @put_store@ to
update the store and reify the result into the current-stage monadic value.
\begin{lstlisting}
  def emit_ap_clo(fun: Rep[Value], arg: Rep[Value], σ: Rep[Store]): Rep[ValSt]
  def ap_clo(ev: EvalFun)(fun: Rep[Value], arg: Rep[Value]): Ans = for {
    σ  <- get_store
    vs <- lift[ValSt](emit_ap_clo(fun, arg, σ))
    _  <- put_store(vs._2)
  } yield vs._1
\end{lstlisting}

Compared with the unstaged version, we can observe a shift of where we use the
evaluation function @ev@: in the unstaged interpreter, the invocation of @ev@
happens in @ap_clo@, i.e., at the application time; while in the staged
interpreter, we eagerly apply @ev@ when denoting the lambda terms to our
next-stage value domains, i.e., at the value-representation time.

\subsection{A Little Bit of Code Generation}

In the previous section, we showed the types of @emit_compiled_clo@ and
@emit_ap_clo@ without concrete implementations. In this section, we briefly
discuss the IR nodes generated by them and sketch the code generation.
We define the IR nodes @IRCompiledClo@ and @IRApClo@ using @case class@
extending from @Def[T]@. @Def[T]@ is a built-in type in the LMS framework,
representing next-stage value definitions of type @T@.
\begin{lstlisting}
  case class IRCompiledClo(f: Rep[(ValSt) => ValSt], λ: Lam, ρ: Rep[Env]) extends Def[Value]
  case class IRApClo(rator: Rep[Value], rand: Rep[Value], σ: Rep[Store]) extends Def[ValSt]
\end{lstlisting}

The IR nodes are manipulated by the LMS passes and finally generated as
next-stage Scala code. To generate code, LMS provides an @emitNode@ method for
programmers to control over what code is generated for each kind of nodes. Here,
we match @IRCompiledClo@ and @IRApClo@, and generate definitions via
@emitValDef@ for them.  For compiled closures, we put the function @f@ and its
accompanying syntactic term and environment to their next-stage value
representation @CompiledClo@. For applications, we know the fact that @fun@ is
an instance of @CompiledClo@ at the next stage, and it has a callable field
@f@. Hence, we generate a next-stage function application @f(arg, σ)@.  The
invocations of @quote@ are used to generate proper names for the other
next-stage values.
\begin{lstlisting}
override def emitNode(sym: Sym[Any], rhs: Def[Any]) = rhs match {
  case IRCompiledClo(f, λ, ρ) => emitValDef(sym, s"CompiledClo($\textdollar${quote(f)}, $\textdollar${quote(λ)}, $\textdollar${quote(ρ)})")
  case IRApClo(rator, rand, σ) => emitValDef(sym, s"$\textdollar$rator.f($\textdollar${quote(rand)}, $\textdollar${quote(σ)})")
  ...
}
\end{lstlisting}