### 2019-09-06 paper/40_absinterp.tex

\section{From Interpreters to Abstract Interpreters} \label{unstaged_abs}

After seeing the unstaged and staged concrete interpreter, now we turn our
focus to abstract interpreters under the same framework. We first present a
lattice representation with binding-time abstraction, as well as simple
abstract domains, such as power set lattice.  Then we show the abstract
components, namely abstract values, abstract environments, and abstract stores.
The abstract interpreter we construct in this section is similar to
\citet{DBLP:journals/pacmpl/DaraisLNH17}'s. For simplicity, it is
context/path/flow-insensitive by our choice. In Section \ref{cfa}, we will
instantiate it to a commonly-used context-sensitive analysis;
\citet{Darais:2015:GTM:2814270.2814308} show how to achieve path- and
flow-sensitivity by varying the monads, which is also applicable to our
unstaged or staged abstract interpreter.

\subsection{Stage Polymorphic Lattices} \label{stagedpoly_lat}

The abstract domains operated by abstract interpreters usually can be
formulated by complete lattices.
A complete lattice over set $S$ is a tuple $\langle S, \sqsubseteq, \top,
\bot, \sqcup, \sqcap \rangle$, where $\top$ is the top element, $\bot$ is the
bottom element, $\sqsubseteq : S \to S \to Bool$ is the
ordering relation, $\sqcup: S \to S \to S$ is the join (least upper bound)
operator, and $\sqcap: S \to S \to S$ is the meet (greatest lower bound)
operator. The trait @SPLattice@ (Figure \ref{fig:splattice}) defines a type
class of stage polymorphic lattices. Similar to @MonadOps@ in Section
\ref{generic_if}, we introduce an additional higher-kinded type @R[_]@ to
annotate the binding-times, where @R[_]@ wraps the data type @S@ and @Boolean@
in these operations.

\begin{figure}[h!]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
  \begin{lstlisting}[style=small]
trait SPLattice[S, R[_]] {
  val ⊤: R[S];  val ⊥: R[S]
  def ⊑(l1: R[S], l2: R[S]): R[Boolean]
  def ⊔(l1: R[S], l2: R[S]): R[S]
  def ⊓(l1: R[S], l2: R[S]): R[S]
}
trait Lattice[S] extends SPLattice[S, NoRep]
  \end{lstlisting}
  \caption{trait \texttt{SPLattice} and \texttt{Lattice}} \label{fig:splattice}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{0.6\textwidth}
\begin{lstlisting}[style=small]
def SetLattice[T]: Lattice[Set[T]] = new Lattice[Set[T]] {
  lazy val ⊤: Set[T] = error("No representation for ⊤")
  lazy val ⊥: Set[T] = Set[T]()
  def ⊑(l1: Set[T], l2: Set[T]): Boolean = l1 subsetOf  l2
  def ⊔(l1: Set[T], l2: Set[T]): Set[T]  = l1 union     l2
  def ⊓(l1: Set[T], l2: Set[T]): Set[T]  = l1 intersect l2
}
\end{lstlisting}
  \caption{The power set instance for lattice} \label{fig:powerset}
\end{subfigure}
\end{figure}

In this section, we simply instantiate @SPLattice@ with the flat binding-time type
@R[T] = T@ (i.e., @NoRep@).
An example of lattices we used in the rest of the paper is the power set
abstract domain (shown in Figure \ref{fig:powerset}): the ordering is the subset
relation, the join is the union operation, and the meet is the intersect
operation. The bottom element of a power set is empty set, and we do not have a
representation of top element for power set.  Other lattices used in the paper, such as
products and maps, can be implemented similarly or by lifting the existing
lattices element-wise or point-wise. Non-relational numerical abstract domains
such intervals can also be implemented in a stage polymorphic way.

\subsection{Abstract Semantics}

In this section, we follow the \textit{abstracting definitional interpreters}
(ADI) approach \cite{DBLP:journals/pacmpl/DaraisLNH17} to refactor our concrete
interpreter to an abstract interpreter.

\paragraph{Abstract Components}

We first widen the concrete values to @AbsValue@. There are two variants of
@AbsValue@: numbers are lifted to a singleton object @IntTop@, which represents
the set of all numbers; closures remain the same. Then the type @Value@ is
defined as a set of @AbsValue@.  The address space is constrained to be finite
to ensure the computability of analysis -- for a simple monovariant analysis,
we directly use variable names for addresses. After defining the @Value@ and
@Addr@ type, the types of environments and stores are automatically lifted to
abstract environments and abstract stores, i.e., @Map[Ident, Addr]@ and
@Map[Addr, Set[AbsValue]]@, respectively.  
\begin{lstlisting}
  trait AbstractComponents extends Semantics {
    sealed trait AbsValue
    case object IntTop extends AbsValue
    case class CloV(lam: Lam, env: Env) extends AbsValue
    type Value = Set[AbsValue]; case class Addr(x: String)
    ...
  }
\end{lstlisting}

It is important that the abstract stores now map addresses to sets of abstract
values, indicating that an address may point to an over-approximated set of
runtime values.  This can be justified by the approximation nature of abstract
interpretation and nondeterminism in the abstract interpreter.  First, when
analyzing a conditional expression, we may not have enough information to
decide which branch will be taken, thus a sound treatment is to explore the both
branches. Also, at some later point, we need to compute the join of two path.
Another nondeterminism comes from function applications, for instance, @f(a)@,
the possible callees ($\lambda$ terms) of @f@ is computed approximately.  As a
result, we may retrieve a set of closures when dereferencing the address of @f@
and we need to apply all of them.

To ensure the abstract interpreter terminates on all programs when computing
the fixed-point, the ADI approach uses a co-inductive mechanism consisting of
two caches that remember the input and output of the @eval@ function.
Here, we first provide the necessary type definitions, and describe
the algorithm later. A @Cache@ is a mapping from configurations @Config@ to
sets of value-store pairs, where the configuration is a triple of the current
expression being evaluated, environment and store. Intuitively, a cache memoizes
the result values and stores for a given program configuration.
\begin{lstlisting}
  type Config = (Expr, Env, Store); type Cache = Map[Config, Set[(Value, Store)]]
\end{lstlisting}

\paragraph{Monads for Abstract Interpretation}
Compared with the concrete interpreter that uses reader and state effects, the
abstract interpreter further introduces the nondeterminism effect and another
reader and state effect for the two caches. The nondeterminism effect is
represented by the @Set[M[_], A]@ monad transformer, where @M@ is the inner
monad type being transformed, and @A@ is the type of elements in the set. We use a
@ReaderT@ for one cache that is not changed during one fixed-point
iteration, and use a @StateT@ for another cache that will be constantly updated
during the analysis.  \citet{DBLP:journals/pacmpl/DaraisLNH17} discuss
different permutations of the monad stack for abstract interpretation. The
following @AnsM@ type shows the monad stack for abstract interpretation (we use
\hl{light gray} to highlight what have been changed from the monad stack of
concrete interpretation):
\begin{lstlisting}[escapechar=!]
trait AbstractSemantics extends AbstractComponents {
  type R[T] = T
  type AnsM[T] = ReaderT[StateT[!\hl{SetT}![!\hl{ReaderT}![!\hl{StateT}![IdM, !\hl{Cache}!, ?], !\hl{Cache}!, ?], ?], Store, ?], Env, T]
  ...
}
\end{lstlisting}

Similar to the concrete scenario, we sketch the @flatMap@ implementation of the
@SetT@ transformer and omit the rest. The field @run@ encapsulated by the monad
is of type @M[SetT[A]]@, where @M[_]@ is the inner monad, and @A@ is the element
type of the set. The @flatMap@ method takes a function that transforms each
element of type @A@ into a @SetT[M, B]@ value, and then folds
all the transformed values into a single monadic value of type @Set[M, B]@.
\begin{lstlisting}
  case class SetT[M[_]: Monad, A](run: M[Set[A]]) {
    def flatMap[B](f: A => SetT[M, B]): SetT[M, B] =
      SetT(Monad[M].flatMap(run) { (s: Set[A]) =>
        s.foldLeft(SetT.empty[M, B])((acc, a) => acc ++ f(a)).run
      }); ... }
\end{lstlisting}

Such monad stack scheme leads to a different type of grounded result. In the concrete
setting, we only have a reader monad (for environments) and a state monad (for
stores). The reader effect is not persistent to the final result. Together
with the value produced by the interpreter, the final result type is a pair of
@Value@ and @Store@. However, under abstract interpretation, we have a @SetT@
inside of the environment monad and store monad. The @Set@ type becomes the
container type of the pairs of values and stores, i.e., @Set[(Value, Store)]@.
Also, note that the reader and state monad for the caches inhibits internally
of the nondeterminism monad; therefore, the final result type is paired
with a cache type from the state monad.
\begin{lstlisting}
  type Result = (R[Set[(Value, Store)]], R[Cache])
\end{lstlisting}

\paragraph{Primitive Operations} The primitive operations are changed according to
the new monad stack scheme and value domains. One of the notable changes is
that the store update operator merges the new values with existing values.
As we mentioned before, sometimes we have to explore the both branches for
conditionals: in method @br0@, we combine the results using the @mplus@
operation from @MonadPlus@, which requires the value domains are
join-semilattices (in our case, @Set@ and @Map@).
\begin{lstlisting}
  def set_store(αv: (Addr, Value)): AnsM[Unit] = liftM(StateTMonad.mod(σ => σ ⊔ Map(αv)))
  def br0(test: Value, thn: => Ans, els: => Ans): Ans = ReaderTMonadPlus.mplus(thn, els)
\end{lstlisting}

The value representation of lambda terms is still a defunctionalized closure
@CloV@, but we lift it to a singleton set.
\begin{lstlisting}
  def close(ev: Expr => Ans)(λ: Lam, ρ: Env): Value = Set(CloV(λ, ρ))
\end{lstlisting}

For function applications, @ap_clo@ looks the same as the concrete interpreter.
The difference is that now the first argument @funs@ is a set of closures (remember
that type @Value@ is an alias of @Set[AbsValue]@), so we can not directly
destruct it to syntactic @Lam@ by pattern matching. Instead, we use an
auxiliary function @lift_nd@ that takes a set and lifts the elements in the set
into the monad stack. Then we can straightforwardly implement the @ap_clo@ in
monadic style, where the closures come from the @SetT@ and thus the
nondeterminism can be naturally handled. The light gray code shows what is
added from the concrete
@ap_clo@.
\begin{lstlisting}[escapechar=!]
  def lift_nd[T](vs: Set[T]): AnsM[T]
  def ap_clo(ev: Expr => Ans)(funs: Value, rand: Value): Ans = for {
    !\hl{CloV(Lam(x, e), $\rho$) <- lift\_nd[AbsValue](funs)}!
    α <- alloc(x)
    _ <- set_store(α → funs)
    rt <- local_env(ev(e))(ρ + (x → α))
  } yield rt
\end{lstlisting}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\paragraph{Caching and Fixpoint Iteration}

As we mentioned earlier, the ADI approach uses a two-cache mechanism to
compute the least fixed-point and prevent non-termination.
The caching algorithm is also called \textit{co-inductive} caching or
\textit{truncated depth-first evaluation} \cite{Rosendahl:AbsIntPL}. It has
been used in other abstract interpreter or fixed point computation
\cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800,
  Rosendahl:AbsIntPL}. The idea is to use a @in@ cache and a @out@ cache during
the depth-first evaluation. The @in@ cache stores the result from the last
iteration, and @out@ cache is used for updating in the current iteration. In the next
iteration, the last @out@ cache will be served as @in@ cache, and an empty cache
is plugged-in to the @out@ slot. Once the @out@ cache does not contain anything
new than the @in@ cache, the fixed-point is reached.

In our monad stack, the @in@ and @out@ caches are modeled by reader effects and
state effects, respectively. We first define several methods to manipulate
the two caches through the monad stack (the implementations are elided).
\begin{lstlisting}
  def ask_in_cache: AnsM[Cache]; def get_out_cache: AnsM[Cache]
  def put_out_cache(out: R[Cache]): AnsM[Unit]
  def set_out_cache(cfg: R[Config], vs: R[(Value, Store)]): AnsM[Unit]
\end{lstlisting}

The co-inductive caching algorithm is implemented as an instrumentation over the
@eval@ function (Figure \ref{fig:coind_cache}), and it also closes the open recursion.
The instrumentation works as follows. Initially, the two caches are both empty.
During the iteration, we first check whether @out@ contains the configuration
@cfg@, representing the current desired computation. If it does, then the result
is directly returned through the monad stack.
Otherwise, we first retrieve the result from @in@ (@⊥@ if @in@ not contains
@cfg@), and update the result into the @out@ cache in the fashion of join.
Then, we invoke @ev@ to compute the result for this iteration, where @ev@ takes
@fix(ev)@ as a self reference.
After the evaluation, the store @σ@ may have been changed. Therefore, 
we get the latest store and construct the result @(v, $\sigma$)@; the
result will be used to update the out cache.

The iteration terminates when the resulting @out@ cache is equivalent to the
input @in@ cache, which indicates that there is no more fact can be discovered.
Therefore, the iteration should end, and we have reached the least fixed point.

\begin{figure}[h!]
  \centering
\begin{lstlisting}
  def fix(ev: (Expr => Ans) => (Expr => Ans)): Expr => Ans = e => for {
    ρ <- ask_env; σ <- get_store; in <- ask_in_cache; out <- get_out_cache; val cfg = (e, ρ, σ)
    rt <- if (out.contains(cfg)) for { // ask if out already contains the desired result
            (v, σ) <- lift_nd[(Value, Store)](out(cfg));  _ <- put_store(σ)
          } yield v
          else for {
            _ <- put_out_cache(out + (cfg → in.getOrElse(cfg, ⊥)))
            v <- ev(fix(ev))(e);  σ <- get_store;  _ <- update_out_cache(cfg, (v, σ))
          } yield v
  } yield rt
\end{lstlisting}
\vspace{-1em}
\caption{The unstaged co-inductive caching algorithm.}
\label{fig:coind_cache}
\vspace{-1em}
\end{figure}

\begin{lstlisting}
  def run(e: Expr): Result = fix(eval)(e)(ρ0)(σ0)(cache0)(cache0).run
\end{lstlisting}

Finally, we override the top-level @run@ method by running the monadic value
with an empty abstract environment $\rho_0$, initial abstract store $\sigma_0$,
and initial caches @cache@$_0$.