### 2019-09-16 paper/90_conc.tex

\section{Conclusion}

We present a simple and elegant framework that extends the first Futamura
projection to abstract interpreters.
To realize it, we design a generic interpreter that abstracts over binding
times and value domains. By instantiating the generic interpreter with the
staged monads for abstract semantics, the staged abstract interpreter can be
readily derived.  The meta-programming approach brings new insights and
implementation techniques to optimize static analyzers via compilation.
Optimizations can be achieved orthogonally to the interpreter and abstract
semantics.  In our evaluation, staging improves the performance of running time
by an order of magnitude.